%% ============================================================================
%% main.tex — Chapter 3 manuscript for the Springer / Elsevier submission.
%%
%% Target venue: primary EMSE (Springer Nature Empirical Software Engineering);
%%               secondary IEEE TSE.
%%
%% Build:
%%     make            # pdflatex + bibtex + pdflatex + pdflatex
%%     make clean      # remove build artefacts
%%     make watch      # continuous rebuild via latexmk
%%
%% Structure (all inputs assembled from paper_draft/):
%%     §1  Introduction                      section_1_introduction.tex
%%     §2  Related Work                      section_2_related_work.tex
%%     §3  Methods                           section_3_methods.tex
%%     §4  Results                           section_4_results.tex
%%     §5  Discussion                        section_5_discussion.tex
%%     §6  Threats to Validity               section_6_threats.tex
%%     §7  Conclusion                        section_7_conclusion.tex
%%     Equations + Algorithms (drop-in)      equations_and_algorithms.tex
%%     Author matter (acks, funding, ethics) section_author_matter.tex
%% ============================================================================

\documentclass[10pt,a4paper]{article}

%% -- Fonts, encoding, geometry -----------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=25mm]{geometry}
\usepackage{microtype}

%% -- Math -------------------------------------------------------------------
\usepackage{amsmath, amssymb, amsthm}
\usepackage{dsfont}         % for \mathds{1}

%% -- Algorithms -------------------------------------------------------------
\usepackage{algorithm}
\usepackage{algpseudocode}

%% -- Figures + tables -------------------------------------------------------
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{caption}
\captionsetup{font=small, labelfont=bf}

%% -- Hyperref (load last of the URL/link packages) ---------------------------
\usepackage[hidelinks]{hyperref}
\usepackage{url}
\usepackage{cleveref}       % \Cref{eq:...}, \Cref{fig:...}

%% -- Bibliography -----------------------------------------------------------
\usepackage[numbers,sort&compress]{natbib}
\bibliographystyle{plainnat}

%% -- Placeholder command for TODO fill-ins -----------------------------------
\newcommand{\PLACEHOLDER}[1]{\textcolor{red!70!black}{\textbf{[TODO: #1]}}}
\usepackage{xcolor}

%% ============================================================================
%% Title block
%% ============================================================================

\title{%
Heterogeneous Multi-SLM Closure of the Docstring--Test--Code Triangle:\\
A Mutation-Testing Study
}

\author{%
Balaji Venktesh V.\thanks{Corresponding author. Email: \texttt{balaji23610040@snuchennai.edu.in}}\\
\PLACEHOLDER{Amsaprabhaa M.} \\
\PLACEHOLDER{Gireesh Sundaram}\\[0.5em]
\normalfont Department of Computer Science and Engineering,\\
\normalfont Shiv Nadar University, Chennai, Tamil Nadu, India
}

\date{\today}

\begin{document}

\maketitle

%% ============================================================================
%% Abstract
%% ============================================================================

\begin{abstract}
Practitioners deploying multi-stage Small-Language-Model (SLM) pipelines for
software engineering face an unanswered empirical question: does it matter
which SLM sits at which stage of the pipeline? We report the first cross-LLM
$\times$ cross-stage evaluation of round-trip closure on the docstring--test--code
triangle, evaluated under a pre-registered 20-cell design of experiments
spanning six open-weight SLMs (all under 30~B parameters) traversing three
closure paths on 150 HumanEval and MBPP functions. Closure is measured under
a strict two-signal validity policy combining an SE-validated automated metric
(mutation kill rate, reference test pass rate, BERTScore F1) with an
external judge SLM's semantic-equivalence rating (DeepSeek-R1). Five findings
emerge. First, the pre-registered H1 composition (phi-4 spec + qwen-coder
test/code) dominates four of five mutation operators empirically, matching
its pre-registered per-operator prediction from a companion mutation-testing
study. Second, phi-4 as the test-generation stage produces filter-failing
tests on approximately half of samples across three independent cells, a
cross-stage pathology invisible in single-stage evaluations. Third, BERTScore
F1 is statistically uncorrelated ($r = 0.02$, $p = 0.359$, $n = 1{,}791$)
with the judge SLM's semantic-equivalence rating on the docstring--docstring
closure path, and the disagreement flips direction across benchmarks in a
signature characteristic of a surface-form similarity signal. Fourth, the
three closure paths carry non-redundant information about cell performance
($F(34, 4865) = 6.21$, $p < 0.001$, partial $\eta^2 = 0.042$). Fifth,
heterogeneity is not free: well-composed hetero cells preserve $78$--$91\%$
of the strongest mono baseline's capability, while wrong-model-at-wrong-stage
cells sacrifice materially. We derive three design guidelines for practitioners
and release an open replication package including 51 unit tests covering the
closure-validity decision and test-filter validity gate.
\end{abstract}

\vspace{1em}
\noindent\textbf{Keywords:} unit test generation; mutation testing;
retrieval-augmented generation; Small Language Models; software engineering;
docstring generation; multi-agent architectures; LLM-as-judge; empirical software
engineering; open-weight language models.

\vspace{1em}
\hrule
\vspace{1em}

%% ============================================================================
%% Body sections
%% ============================================================================

\input{section_1_introduction.tex}
\input{section_2_related_work.tex}

%% Section 3 references equations + algorithms defined in the block below.
%% Include the equations/algorithms file INSIDE section_3_methods.tex once
%% the labels are stable; for now, place them here so \label targets exist
%% before §4 and §5 reference them.
\input{equations_and_algorithms.tex}

\input{section_3_methods.tex}
\input{section_4_results.tex}
\input{section_5_discussion.tex}
\input{section_6_threats.tex}
\input{section_7_conclusion.tex}

%% ============================================================================
%% Author matter (acknowledgements, funding, ethics, etc.)
%% ============================================================================

\input{section_author_matter.tex}

%% ============================================================================
%% Bibliography
%% ============================================================================

\bibliography{references}

\end{document}
