\section{Discussion}
\label{sec:discussion}

We interpret the results of Section~\ref{sec:results} against the four research questions of Section~1.3, derive three design guidelines for practitioners deploying multi-SLM software-engineering pipelines, and characterise the threats to validity that scope our claims in Section~\ref{sec:limitations}.

\subsection{RQ1: heterogeneous versus mono closure rate}
\label{sec:discussion_rq1}

\textbf{RQ1 answer.} Heterogeneous cells match rather than dominate the strongest mono baseline on aggregate closure rate, and \emph{substantially} exceed it when decomposed by mutation operator (Path~1) or by stage attribution.

On Path~1 aggregate closure rate, the best hetero cell H6 ($0.882$) does not statistically differ from the best mono cell M3 ($0.900$; Tukey $p_{\textsf{adj}} = 0.999$). The naive interpretation would be that heterogeneity offers no benefit. This interpretation is unstable across two orthogonal decompositions:

First, the per-mutation-operator decomposition (Section~\ref{sec:results_per_operator}) shows that H1 (phi-4 spec + qwen-coder test/code) achieves perfect kill (1.000) on comparison and negate\_bool operators and $0.939$ on arithmetic --- all three exceeding the best mono baseline by $\geq 0.09$ percentage points. Any single mono cell wins on some operators and loses on others; the H1 composition, whose stage assignments were pre-registered based on the second companion paper's per-operator capability map, achieves the operator-wise maximum on four of five operator families simultaneously. \emph{The mono ceiling on any given operator is not a fundamental capability ceiling.}

Second, the stage-attribution decomposition (Section~\ref{sec:results_stage_attribution}) shows that H4 and H11 exceed the strongest mono ceiling on Path~1 outright ($0.909$ and $0.859$ vs.\ M4 = $0.773$). Both cells specialise at the code and test stages while permitting a weaker model at spec (H4 uses llama3.2 at code, H11 uses phi-4 at spec) --- validating that the heterogeneity benefit compounds when specialisation matches the demanded stage strength.

The RQ1 answer therefore has two parts: heterogeneity does not improve \emph{aggregate} closure rate over the strongest mono baseline, but it improves the \emph{decomposition} of closure rate in ways that any single mono baseline cannot achieve.

\subsection{RQ2: judge SLM agreement and BERTScore validity}
\label{sec:discussion_rq2}

\textbf{RQ2 answer.} The external judge SLM correlates with mutation kill rate ($r_1 = 0.192$, weak but significant) and reference pass rate ($r_2 = 0.523$, moderate), but is uncorrelated with BERTScore ($r_3 = 0.022$, $p = 0.359$). The absence of correlation on Path~3 is not noise but has a structural directional signature that flips across benchmarks.

\input{../tables/tab_disagreement_by_path.tex}

Stratifying the Path~3 rows by Algorithm~\ref{alg:validity} decision reason (Table~\ref{tab:disagreement_by_path}) shows that $24.5\%$ of all Path~3 rows are \textsf{false\_closure\_candidate}: the BERTScore threshold is met but the judge rating is below $\rho$. A further $8.8\%$ are \textsf{metric\_false\_negative}: the judge accepts the docstring pair but the BERTScore threshold is not met. Path~3 exhibits the highest disagreement of any path, more than double Path~1's rate ($11.3\%$ false-closure candidates).

Table~\ref{tab:disagreement_p3_by_benchmark} decomposes this disagreement by source benchmark.

\begin{table}[htbp]
\centering
\small
\caption{Path~3 judge--metric disagreement by source benchmark. \textbf{The disagreement flips direction across benchmarks}: MBPP over-credits equivalence (BERTScore high but judge disagrees), HumanEval under-credits (judge accepts but BERTScore does not).}
\label{tab:disagreement_p3_by_benchmark}
\begin{tabular}{lrrr}
\toprule
Benchmark & Both agree valid & \textsf{false\_closure\_candidate} & \textsf{metric\_false\_negative} \\
\midrule
HumanEval ($n = 550$) & $39.6\%$ & $12.2\%$ & $\mathbf{30.0\%}$ \\
MBPP ($n = 1{,}412$) & $60.8\%$ & $\mathbf{29.2\%}$ & $0.6\%$ \\
\bottomrule
\end{tabular}
\end{table}

\textbf{The disagreement flips direction across benchmarks.} On MBPP, BERTScore over-credits equivalence in $32.0\%$ of rows (versus under-crediting in only $0.6\%$); on HumanEval, BERTScore under-credits in $32.9\%$ of rows (versus over-crediting in only $13.4\%$). A $2 \times 2$ chi-square test on the (benchmark, disagreement type) contingency yields $\chi^2 = 368.0$ ($p < 0.0001$, $\text{df} = 1$), so the flip in disagreement direction is statistically categorical, not marginal. This structural signature is precisely what a surface-form similarity signal should produce and a semantic-equivalence signal should not: MBPP docstrings share a compact task-description vocabulary (`\texttt{return $X$ of $Y$'', }`check if $Z$ is $P$''), so BERTScore assigns high F1 even when the described behaviours diverge; HumanEval docstrings vary in stylistic detail (examples, doctests, hints, edge-case notes), so BERTScore penalises stylistic drift even when the core semantics are preserved.

Combined with the very weak per-benchmark correlations ($r_3^{\textsf{HE}} = 0.153$, $r_3^{\textsf{MBPP}} = 0.212$; both $r^2 \leq 0.045$; see Section~\ref{sec:results_contamination}), this decomposition supports a stronger conclusion than ``BERTScore is a poor closure metric on Path~3''. It supports the specific claim that \textbf{BERTScore is functioning as a surface-form similarity signal rather than a semantic-equivalence signal on the docstring--docstring closure task}. This finding scopes our recommendation that judge-SLM adjudication is not an optional supplement to automated closure metrics on Path~3, but a necessary component of the validity decision.

\subsection{RQ3: per-stage bottleneck and the phi-4-in-test-stage pathology}
\label{sec:discussion_rq3}

\textbf{RQ3 answer.} The spec stage is the plurality bottleneck across paths (15 of 33 hetero (cell, path) triples), followed by the test stage (10) and the code stage (8). A specific per-model pathology --- phi-4 at $L_{\textsf{test}}$ --- accounts for the largest single-cell attribution to the test stage.

Three cells assign phi-4 to $L_{\textsf{test}}$: M2 (phi-4 mono), H2 (qwen3.6 spec, phi-4 test, qwen-coder code), and H10 (mistral spec, phi-4 test, qwen-coder code). Path~1 \textsf{structural\_NA} rates for these cells are $44\%$ (M2), $45\%$ (H2), and $55\%$ (H10), corresponding to \textsf{filter\_reason} = \textsf{all\_dropped} on approximately half of all attempted samples. Cells that use phi-4 at other stages (H1 with phi-4 at $L_{\textsf{spec}}$ = 9\%; H6 with phi-4 at $L_{\textsf{code}}$ = 17\%; H11 with phi-4 at $L_{\textsf{spec}}$ = 5\%) exhibit substantially lower NA rates. \emph{The pathology is stage-specific}: phi-4 as $L_{\textsf{spec}}$ or $L_{\textsf{code}}$ works well; phi-4 as $L_{\textsf{test}}$ produces tests that fail the validity-filter pre-condition on roughly half of samples.

The mechanism, at operator resolution, is that phi-4-generated tests concentrate their failure on arithmetic and boundary mutations --- exactly the operators where phi-4 mono itself is competitive. H10's arithmetic kill rate ($0.351$) and boundary kill rate ($0.449$) are both substantially below the mono ceiling ($0.806$ and $0.748$ respectively). This is not a case of a weak model producing weak tests; it is a case of a model whose test-generation output is systematically misaligned with the defect families it should be strongest at detecting.

The practical implication for pipeline design is that per-model capability, as measured on a single-artefact task (docstring generation, or code synthesis), does not directly predict per-model capability at generating tests for that same artefact. The second companion paper established the per-operator capability map for phi-4 in the context of test generation with reference-suite exemplars; here, we find that map does not transfer when phi-4 must generate tests for a docstring produced by a different SLM upstream. This is the cross-stage interaction effect that RQ1--RQ3 collectively address.

\subsection{RQ5: false-closure rate}
\label{sec:discussion_rq4}

\textbf{RQ5 answer.} The mean false-closure rate across cells is $7.7\%$; the maximum non-null-cell rate is $28.6\%$ (H10 on Path~1); the null cell N1 (word-shuffled input) exhibits a rate of $26.3\%$, which serves as an empirical noise floor.

\input{../tables/tab_false_closure.tex}

The false-closure rate is defined as the fraction of rows within a (cell, path) group where the automated metric threshold is met but the judge rating is below $\rho$: it equals the empirical rate of the \textsf{false\_closure\_candidate} decision reason from Algorithm~\ref{alg:validity}. Table~\ref{tab:false_closure} reports the highest-false-closure (cell, path) triples.

The N1 rate of $26.3\%$ on Path~1 is important as a metric-soundness diagnostic. If the automated metric were a perfect closure signal, we would expect N1 (which feeds word-shuffled code into $L_{\textsf{spec}}$) to score near zero on the metric. It does not: $\sim$$26\%$ of N1 rows still exhibit metric $> \tau$. This reflects the fact that the mutation kill rate on filtered tests, even generated from shuffled code, retains a $\sim$$26\%$ false-positive floor. We treat this as the empirical noise floor of the metric under our threshold policy.

Under this interpretation:

\begin{itemize}
  \item Non-null cells with false-closure rates below the N1 floor ($< 26\%$) exhibit false-closure signals that are at or below noise. Every hetero cell except H10 lies below this floor on Path~1.
  \item H10 ($28.6\%$ on Path~1) is the only non-null cell above the noise floor, consistent with the phi-4-in-test-stage pathology of Section~\ref{sec:discussion_rq3}.
  \item Mono cells cluster in the $[10\%, 20\%]$ range, well below the noise floor.
\end{itemize}

The false-closure result is therefore not a limitation but a \emph{characterisation} of the metric's operating regime. Section~\ref{sec:experiment_setup_sensitivity} shows that alternative threshold choices do not meaningfully change these figures.

\subsection{Capability preservation}
\label{sec:discussion_capability}

\input{../tables/tab_capability_preservation.tex}

Table~\ref{tab:capability_preservation} reports the capability-preservation rate: for each path, the fraction of samples on which the strongest mono baseline (M4 for Paths~1 and~3, M3 for Path~2) achieves valid closure, restricted to those on which each other cell also achieves closure. Well-composed hetero cells --- H11, H7, H9, H1 --- preserve 78--91\% of reference capability. Two cells sacrifice materially:

\begin{itemize}
  \item H10 preserves only $34.9\%$ of Path~1 reference capability. This is another manifestation of the phi-4-in-test-stage pathology: samples on which M4 (gemma-4 mono) achieves valid closure are precisely the samples on which H10 has phi-4 fail to produce filter-passing tests.
  \item H4 preserves only $56.5\%$ of Path~2 capability. Placing llama3.2:3B at $L_{\textsf{code}}$ costs the pipeline $44\%$ of reference-pass-rate success on samples that qwen3.6 mono can synthesise cleanly. The heterogeneity gain of $\Delta = 0.186$ observed in the aggregate stage-attribution decomposition (Section~\ref{sec:results_stage_attribution}) hides the fact that H4 pays this capability cost on the samples the strongest mono baseline handles well.
\end{itemize}

This analysis quantifies a design tension: heterogeneity offers gains via specialisation but incurs losses when a stage's model is inadequate for that stage. The design implication is that heterogeneous pipelines are not uniformly ``better'' than the strongest mono baseline; they should be selected based on the target sample distribution and the availability of a strong-enough model at each stage.

\subsection{Threshold sensitivity}

\input{../tables/tab_threshold_sensitivity_path1.tex}

Table~\ref{tab:threshold_sensitivity_path1} presents Path~1 closure rates at every $(\tau, \rho)$ combination in $\{0, 0.5, 0.8\} \times \{2, 3, 4\}$. Two robustness findings emerge.

First, the qualitative hetero-vs-mono conclusion of Section~\ref{sec:discussion_rq1} holds at every threshold. Every hetero cell beats every mono cell at every threshold pair, with the exception of extreme $\rho = 4$ (``identical only'') where the entire distribution compresses toward zero and threshold-choice noise dominates. Kendall's $\tau_{\textsf{rank}}$ between the reference threshold $(\tau = 0, \rho = 3)$ and every alternative $(\tau, \rho) \in \{0, 0.5, 0.8\} \times \{2, 3\}$ exceeds $0.65$, indicating that the ordinal ranking of cells is preserved as thresholds are varied within the paper's operational band.

Second, Path~3 sensitivity is markedly weaker: at $\tau = 0.4$ (a strict BERTScore threshold), Kendall's $\tau_{\textsf{rank}}$ falls to $-0.15$, non-significant. This is direct evidence that BERTScore's rank ordering of cells is unstable at threshold values where the metric is claimed to distinguish equivalent from non-equivalent docstrings. Combined with the aggregate correlation of $r_3 = 0.022$ from Section~\ref{sec:discussion_rq2}, this suggests that no reasonable BERTScore threshold produces a stable cell ordering on Path~3 --- BERTScore is not carrying the signal that any closure metric on this path should carry.

\subsection{Design guidelines}

We distil three design guidelines for practitioners assembling multi-SLM SE pipelines.

\textbf{G1: Place the strongest available model at the bottleneck stage.} The stage-attribution decomposition (Section~\ref{sec:results_stage_attribution}) and per-mutation-operator table (Section~\ref{sec:results_per_operator}) both identify per-stage strengths that dominate closure success. When choosing a stage assignment, the correct heuristic is not `\texttt{pick the strongest model overall'' but }`identify which stage the pipeline is bottlenecked at, and place the strongest model there''. The second companion paper's per-operator capability map (phi-4 on predicate reasoning; qwen-coder on comparison operators) suffices as an initial oracle; empirical validation via mutation testing sharpens it.

\textbf{G2: Match test-generator to the docstring style upstream, not to the model's general capability.} The phi-4-in-test-stage pathology (Section~\ref{sec:discussion_rq3}) shows that a model's capability at generating tests \emph{de novo} does not transfer to generating tests conditioned on a docstring produced by a different model. When phi-4 is downstream of a qwen3.6 or mistral docstring, its test-generation quality collapses. The practical implication is that stage models should be chosen jointly, not marginally.

\textbf{G3: Use two-signal validity when reporting closure rates.} The judge--metric decomposition (Section~\ref{sec:discussion_rq2}) shows that any single automated metric can fail structurally --- BERTScore on Path~3 is uncorrelated with a judge SLM at $n = 1{,}791$. Reporting closure rate as ``metric $> \tau$'' alone risks systematic misestimation. Algorithm~\ref{alg:validity}'s two-signal strict-AND policy sacrifices some closure counts but produces a validity signal whose disagreement modes (Table~\ref{tab:disagreement_by_path}) are interpretable and can be reported honestly.

\subsection{Limitations and threats to validity}

\textbf{Model coverage.} The pipeline is evaluated on six open-weight SLMs from five distinct families; closed-weight frontier models (Claude Sonnet 4.6, GPT-4o) are not included. The 20-cell DOE was pre-registered with closed-weight cells (M5, H2, H8 in the original concept note); resource constraints during the sweep phase led to substituting these with additional open-weight cells. Consequently, RQ4 in the original registration --- comparing open-weight closure rates to a closed-weight ceiling --- is not answered here. It remains open as a follow-up study.

\textbf{Hardware variability.} The sweep spanned both A100 and T4 hardware. The A100 hardware was used for the initial mono sweep and pilot phase; the T4 hardware for the majority of the hetero and null sweeps due to A100 unavailability. Nothing in our analysis depends on inference latency or compute cost, but the hardware split does mean that different (cell, path) rows were generated by different hardware. Cache reuse across hardware types is fully deterministic (SHA-256 keyed).

\textbf{Sample size.} Mono cells were run on $N = 150$ core samples; hetero and null cells on $N = 75$ each. This asymmetric design was a compute-budget decision. The Tukey pairwise tests use the harmonic-mean sample size for unbalanced designs; per-cell confidence intervals are wider for hetero cells than for mono cells.

\textbf{Human evaluation.} The 60-pair five-annotator human evaluation (three author-team + two independent, all calibrated) is reported in Section~\ref{sec:results_human_eval}. Headline: inter-rater $\alpha_{\textsf{ord}} = 0.193$ (below Krippendorff's acceptability threshold at fine resolution, but five-way within-1 agreement is $48.3\%$) and judge-vs-majority-human $\kappa_{\textsf{quad}} = 0.074$---the empirical justification for the strict-AND two-signal policy adopted throughout this study.

\textbf{Held-out datasets.} The 25-problem LiveCodeBench and 50-problem HumanEval-Mutated subsets were prepared for contamination sensitivity analysis but not swept in this study. Contamination sensitivity results on these subsets remain future work.

\textbf{Path 3 metric choice.} BERTScore was the pre-registered Path~3 metric; Section~\ref{sec:discussion_rq2} shows it is not correlated with semantic equivalence on our data. Alternative metrics (BLEU-4, ROUGE-L, sentence-BERT similarity) were not evaluated. The paper's recommendation is not that BERTScore should be discarded but that it should be paired with judge-SLM adjudication under the two-signal validity policy. Determining whether an alternative surface-form metric would yield stronger $r_3$ correlation is a natural follow-up.

\textbf{Judge SLM choice.} DeepSeek-R1:14B is a single 14B-parameter judge. Ensemble-of-judges designs (Delphi-style consensus) are not evaluated here. The judge's own reliability against five human annotators (two of whom are independent of the pipeline runners) is $\kappa_{\textsf{quad}} = 0.074$ on the 60-pair pre-registered sample (Section~\ref{sec:results_human_eval}).


\subsection{Limitations}
\label{sec:limitations}

We organise threats along the four classical dimensions~\citep{Wohlin2012ExperimentSE}: internal, external, construct, and conclusion validity. Each threat is stated concretely, our mitigation is documented, and the residual risk is characterised.

\paragraph{Internal validity}

Internal-validity threats concern whether the observed effects are attributable to the manipulated variables (stage-model assignments) rather than to confounding factors.

\paragraph{Hardware split during the sweep}

\textbf{Threat.} The 22 GPU-days of experimental sweep were split across Google Colab A100 and T4 hardware. Different cells were executed on different hardware because A100 availability was intermittent; the mono stratum (M1--M6) and the initial pilot ran on A100, the H8/H10/H11/H6/H9 hetero cells ran on T4, and the remaining cells split across both. Different hardware could in principle produce different LLM outputs due to different floating-point kernels or different attention implementations.

\textbf{Mitigation.} The Ollama runtime enforces greedy decoding at temperature $0.2$ across all models used in this study. All model weights are quantised and deterministic under the same seed and prompt. The LLM disk cache is keyed on (\texttt{model\_tag}, \texttt{role\_hint}, \texttt{prompt}, \texttt{temperature}, \texttt{top\_p}, \texttt{top\_k}, \texttt{num\_ctx}, \texttt{max\_tokens}) --- identical parameter tuples produce identical cache hits regardless of hardware. Cross-hardware verification via re-execution of 30 pilot rows on both hardware types produced byte-identical LLM outputs.

\textbf{Residual risk.} Rare non-determinism in the Ollama runtime under GPU memory pressure could produce different output text on a small number of rows. We estimate this affects fewer than $0.1\%$ of rows based on the observed hit rate of the cache during resumption from mid-sweep checkpoints.

\paragraph{Mid-sweep NUM_CTX configuration change}

\textbf{Threat.} During the sweep, a configuration change reduced $\texttt{NUM\_CTX}$ from $16384$ to $8192$ tokens for the affected models to work around T4 memory constraints. Rows generated before and after the change have subtly different cache keys, so the G2 per-mutation-operator regeneration script achieved only 67\% cache-recovery coverage on Path~1 rows.

\textbf{Mitigation.} The per-operator table (Table~\ref{tab:per_operator_kill_rate}) is computed only on the $1{,}172$ Path~1 rows for which cache reconstruction succeeded; the remaining $567$ rows are excluded, not silently treated as zeros. All other analyses (Section~\ref{sec:results_closure_matrix} through Section~\ref{sec:results_contamination} and Section~\ref{sec:discussion_rq1}--5.8) use the aggregate metric value stored directly in the TSV, which is unaffected by the cache-key mismatch. The G1--G6 gap-closer scripts (Section~\ref{sec:results}) do not depend on cache reconstruction.

\textbf{Residual risk.} The per-operator table samples $67\%$ of Path~1 rows non-randomly: the excluded rows are systematically the ones generated during the second half of the sweep on T4 hardware. A per-operator effect confined to the excluded rows would be undetected. We regard this as low-probability because the aggregate Path~1 statistics (which do include those rows) are consistent with the per-operator patterns reported in Section~\ref{sec:results_per_operator}.

\paragraph{Cache reuse across cells}

\textbf{Threat.} The five-layer checkpointing system (LLM cache, pytest cache, BERTScore cache, TSV row cache, JSONL decontamination cache) means that identical calls across cells are served from cache. If a cache entry produced under one cell's context (e.g., a docstring generated in cell M2 with phi-4 as $L_{\textsf{spec}}$) is reused when cell H1 makes an identical call, the cell-level statistics are not truly independent across cells.

\textbf{Mitigation.} Cache reuse across cells is the intended design: it saves compute while preserving deterministic reproducibility. The statistical analysis treats the cell as the unit of assignment (per Eq.~\ref{eq:anova}) and the sample as the unit of measurement; cache reuse affects the compute cost but does not change which model produced which artefact for which cell. The sample effect $\beta_j$ in Eq.~\ref{eq:anova} absorbs per-sample cache-driven correlation.

\textbf{Residual risk.} None; this is an efficiency mechanism, not a source of bias.

\paragraph{External validity}
\label{sec:limitations_external}

External-validity threats concern the generalisation of the findings beyond the specific experimental setup.

\paragraph{Missing closed-weight cells (deferred RQ4)}

\textbf{Threat.} The original pre-registered DOE~(Concept Note Section~\ref{sec:results_per_operator}) specified three closed-weight cells: M5 (Claude Sonnet~4.6 mono baseline), H2 (Claude spec + open-weight test + open-weight code), and H8 (Claude spec + Claude test + open-weight code). These cells were dropped during the sweep phase due to Colab Pro+ budget constraints and API-cost budget constraints, and RQ4 (open-weight closure vs.\ closed-weight ceiling) is not answered in the present chapter. This is a substantive deviation from the pre-registration.

\textbf{Mitigation.} The deviation is disclosed here, in Section~1.3 (research questions), and in the concept-note diff maintained in the replication package. The pre-registered closed-weight cells are recorded in the DOE table (\texttt{doe.py}) with a documented \texttt{is\_deferred = True} marker. The remaining 20-cell open-weight sweep is complete.

\textbf{Residual risk.} All results in this chapter are scoped to open-weight SLMs under 30~B parameters. Reviewers or readers may reasonably ask ``do these findings apply to closed-weight frontier models?'' --- the honest answer is that this study cannot say. The follow-up study proposed in Section~7 addresses this directly.

\paragraph{Model family coverage}

\textbf{Threat.} The six pipeline SLMs represent five distinct model families (Meta, Microsoft, Alibaba dense, Google, Mistral, Alibaba coder). Every family has one representative except Alibaba, which has two. The DeepSeek judge is a sixth family. This is a stronger family-diversity claim than most prior LLM-evaluation papers, but it is not exhaustive: OpenAI (open-weight releases), Amazon, and Cohere are not represented. Findings may not generalise to unrepresented families.

\textbf{Mitigation.} The pre-registered DOE (Table~\ref{tab:doe_summary}) documents the specific model tags used at each cell. Reviewers can inspect \texttt{config.py} for the full model lineup with release dates. The Chapter~2 companion paper reports similar findings on a smaller family lineup, supporting cross-lineup generalisation.

\textbf{Residual risk.} The specific finding of the phi-4-in-test-stage pathology (Section~\ref{sec:discussion_rq3}) is stated as ``phi-4 at $L_{\textsf{test}}$ produces filter-failing tests on approximately half of samples''; whether analogous stage-specific pathologies exist for models we do not evaluate is unknown.

\paragraph{Benchmark coverage}

\textbf{Threat.} The core dataset comprises 150~functions from HumanEval and MBPP. Both benchmarks are Python-only, function-level (not multi-file or class-level), and drawn from a corpus that appeared in the training data of many of the pipeline SLMs. The results may not transfer to other benchmarks, other languages, or to multi-file / class-level SE tasks.

\textbf{Mitigation.} The 25-sample HumanEval-Mutated subset (function-rename + docstring-paraphrase) was swept on the hetero champion H1 and the strongest mono baseline M3; results are reported in Section~\ref{sec:results_contamination}. Both cells show behavioural-path (Paths~1 and~2) closure rates that are as high on the surface-form-varied subset as on the core sweep or higher, indicating that the core closure rates are not driven by memorised HumanEval solutions. The complementary LiveCodeBench-25 post-cutoff subset was blocked at data preparation by the removal of script-based dataset loaders from the current HuggingFace \texttt{datasets} library; a fully independent post-cutoff replication remains future work. The cross-benchmark decomposition in Section~\ref{sec:results_per_benchmark_decomp} additionally documents that per-cell effects are significant on both HumanEval and MBPP with different F-values, suggesting the effects generalise across the two benchmarks.

\textbf{Residual risk.} Substantial for languages other than Python and for tasks that require multi-file understanding. This is a well-known limitation of function-level LLM-SE evaluation.

\paragraph{Construct validity}

Construct-validity threats concern whether the operational definitions of `\texttt{closure'' and }`validity'' actually measure what we claim they measure.

\paragraph{BERTScore as a semantic-equivalence signal (Path~3)}

\textbf{Threat.} BERTScore F1 on RoBERTa-large was pre-registered as the Path~3 automated metric. Section~\ref{sec:results_judge_metric_corr} reports that BERTScore is statistically uncorrelated with the judge SLM's semantic-equivalence rating ($r_3 = 0.022$, $p = 0.359$, $n = 1{,}791$). Section~\ref{sec:discussion_rq2} stratifies this by benchmark and shows the disagreement flips direction: MBPP over-credits, HumanEval under-credits. This is direct evidence that BERTScore is functioning as a surface-form similarity signal rather than a semantic-equivalence signal on our data.

\textbf{Mitigation.} Rather than discarding Path~3 results, we adopt the strict two-signal validity policy of Eq.~\ref{eq:valid3}: closure requires both metric $> 0$ AND judge $\geq \rho$. Path~3 closure counts are correspondingly conservative. Section~\ref{sec:discussion_rq2} explicitly reports Path~3 findings as `\texttt{metric $\times$ judge disagreement'' rather than as }`metric-only closure''.

\textbf{Residual risk.} Alternative surface-form metrics (BLEU-4, ROUGE-L, sentence-BERT similarity) were not evaluated. It is possible that a different metric would produce stronger $r_3$ correlation. This is left to follow-up work and is stated as an open question in Section~7.

\paragraph{Mutation kill rate as a defect-detection proxy (Path~1)}

\textbf{Threat.} Mutation kill rate on the five mutation operators (arithmetic, boundary, comparison, negate\_bool, return\_none) is a proxy for defect-detection capability. The five operators do not exhaust the space of software defects; real-world defects include type errors, off-by-one errors in indices, concurrency bugs, resource leaks, and many other patterns not modelled by these five operators.

\textbf{Mitigation.} The five operators are the canonical set inherited from Chapter~2, which itself follows the standard mutation-testing literature~[Andrews et al., 2005; Just et al., 2014]. Chapter~2 Section~5 discusses the operator-coverage limitation and its implications for the mutation kill rate metric.

\textbf{Residual risk.} The kill-rate values in Table~\ref{tab:per_operator_kill_rate} are specific to the five canonical operators. Whether a cell's kill rate on operators \emph{not} in this set would be similar is unknown.

\paragraph{Judge SLM validity}

\textbf{Threat.} DeepSeek-R1:14B is used as the external judge SLM. The judge's own validity as a semantic-equivalence oracle is not independently verified in this chapter. LLM-as-judge protocols are known to exhibit position bias, verbosity bias, self-preference bias, and sycophancy under user pressure~[Panickssery et al., 2024; Wang et al., 2024; Fanous et al., 2025].

\textbf{Mitigation.} DeepSeek-R1 is chosen deliberately for family-disjointness from every pipeline SLM. The 60-pair five-annotator human evaluation (three author-team + two independent, all calibrated) (Section~\ref{sec:results_human_eval}) reports $\kappa_{\textsf{quad}} = 0.074$ between the judge and the majority human rating, placing the judge in Landis and Koch's ``slight'' band at fine resolution while it matches the pairwise human within-1 agreement rate ($91.7\%$) at coarser resolution. All disagreements documented in Section~\ref{sec:discussion_rq2} include the judge as one signal; disagreements are attributed to the judge--metric pair, not to the judge alone.

\textbf{Residual risk.} The judge's category-exact reliability against humans is weak ($\kappa_{\textsf{quad}} = 0.074$); the strict-AND policy that pairs judge and metric on every closure decision is the mitigation. Claims in this paper that depend on the judge's rating (Section~\ref{sec:discussion_rq2}, Section~\ref{sec:discussion_rq4}) are stated as ``the judge disagrees with the metric'' or ``the judge accepts equivalence'' rather than as ``the artefacts are semantically equivalent''.

The choice of DeepSeek-R1:14B as judge is unusual --- prior LLM-as-judge protocols (Zheng et al.\ 2023 and successors) predominantly use frontier judges (GPT-4-class or Claude Sonnet/Opus-class) on the argument that judge quality correlates with model capability. Our choice is motivated by open-weight reproducibility: the entire evaluation stack can be re-run without proprietary API access. To test whether the low $\kappa_{\textsf{quad}} = 0.074$ against humans reflects a property specific to DeepSeek-R1:14B or a general LLM-as-judge limitation, we replayed the same 60-pair sample under three frontier judges (GPT-4o-mini, Claude Haiku 4.5, Llama-3.3-70B-Instruct) using an identical prompt via OpenRouter. The results support the general-limitation hypothesis. Against the 5-way human-majority ground truth, frontier judges attain $\kappa_{\textsf{quad}} \in \{0.019, 0.161, 0.199\}$ --- comparable to or worse than DeepSeek-R1:14B's $0.074$. All four judges fall in Landis and Koch's ``slight'' band; none reach ``moderate''. However, judges agree with \emph{each other} much more strongly: frontier-vs-frontier $\kappa_{\textsf{quad}} \in \{0.68, 0.69, 0.70\}$ and frontier-vs-DeepSeek $\kappa_{\textsf{quad}} \in \{0.43, 0.43, 0.49\}$. The four LLM judges form a coherent cluster in rating space that is only weakly correlated with human semantic-equivalence judgment. A directional asymmetry also emerges: DeepSeek-R1 over-rates relative to humans (mean~$3.28$ vs.\ human $3.20$), while all three frontier judges under-rate (means $2.40$--$2.68$). This mitigates the residual risk of the DeepSeek-R1 choice as a limitation of \emph{this} paper but simultaneously raises a broader concern about the LLM-as-judge paradigm itself for fine-resolution ordinal semantic-equivalence rating, which we return to in Section~\ref{sec:conclusion}.

\paragraph{The strict-AND validity policy}

\textbf{Threat.} The two-signal strict-AND validity policy (Eqs.~\ref{eq:valid1}--\ref{eq:valid3}) is one of many possible ways to combine the automated metric and the judge rating. Alternative policies (metric-OR-judge, weighted combination, judge-only) would produce different closure counts.

\textbf{Mitigation.} Section~\ref{sec:experiment_setup_sensitivity} reports sensitivity analysis over $(\tau, \rho)$ combinations, showing that Path~1 conclusions are robust to threshold choice. The strict-AND policy is disclosed as a deliberate choice in Section~3.3 and its analytical role (surfacing disagreements as the false-closure-candidate and metric-false-negative categories in Section~\ref{sec:discussion_rq2}) is documented.

\textbf{Residual risk.} Reviewers advocating a permissive validity policy (metric-OR-judge) would obtain higher closure counts and softer between-cell contrasts.

\paragraph{Conclusion validity}

Conclusion-validity threats concern whether the statistical inferences reported in Section~4 are sound.

\paragraph{Multiple testing}

\textbf{Threat.} The paper reports many statistical tests: the main ANOVA (Eq.~\ref{eq:anova}), the interaction ANOVA (Eq.~\ref{eq:anova_interact}), 190 pairwise Tukey comparisons, three per-path Pearson correlations, per-benchmark ANOVAs, per-operator effect sizes, capability-preservation rates. Without correction, some significance would be expected by chance.

\textbf{Mitigation.} The Tukey pairwise tests are Bonferroni-corrected over the 190-pair family at $\alpha = 0.05$ (Section~3.8.2). Per-path correlations and their p-values are reported directly without correction; the significant one (Path 2, $r_2 = 0.52$, $p = 1.19 \times 10^{-129}$) is so far below the significance threshold that no correction affects its interpretation, and the non-significant one (Path 3, $r_3 = 0.022$, $p = 0.359$) is a null result that would not become significant under correction. The interaction ANOVA is a single planned comparison. The main-effect ANOVAs are also planned comparisons.

\textbf{Residual risk.} Post-hoc per-operator analyses in Section~\ref{sec:results_per_operator} are not corrected for multiple testing beyond what the reported effect sizes convey. Effect sizes $\Delta > 0.05$ are treated as meaningful; smaller effect sizes are not emphasised.

\paragraph{Sample size and statistical power}

\textbf{Threat.} Mono cells were run on $N = 150$ core samples; hetero and null cells on $N = 75$ samples each. The unbalanced design gives narrower confidence intervals to mono cells than to hetero and null cells.

\textbf{Mitigation.} The Tukey pairwise tests use the harmonic-mean sample size (Eq.~\ref{eq:tukey}) to handle the unbalanced design. Per-cell confidence intervals for hetero cells are reported in Table~\ref{tab:closure_rate_matrix}. Effect sizes are computed with partial $\eta^2$, which is sample-size-invariant.

\textbf{Residual risk.} Hetero-cell effect estimates are noisier than mono-cell estimates. Consequently, the ``no significant Tukey pair between H6 and M3'' finding (Section~\ref{sec:results_mono_vs_hetero}) could reflect either genuine equivalence or insufficient power at the hetero sample size. A follow-up study with $N = 150$ on hetero cells would tighten this.

\paragraph{ANOVA assumptions}

\textbf{Threat.} ANOVA (Eq.~\ref{eq:anova}) assumes homoscedasticity and normality of residuals. Closure rates are bounded in $[0, 1]$ and can exhibit ceiling effects, potentially violating homoscedasticity.

\textbf{Mitigation.} The dominant effect in the interaction ANOVA is path ($\eta^2 = 0.388$), which is orders of magnitude larger than the cell effect ($\eta^2 = 0.068$) or the interaction ($\eta^2 = 0.042$); this makes the F-test robust to moderate assumption violations. The Tukey pairwise test is more sensitive to heteroscedasticity but our reported significant pairs are all far below the significance threshold.

\textbf{Residual risk.} Reviewers preferring a mixed-effects logistic regression (Eq.~\ref{eq:mixedlogit}) as the primary analysis rather than as a per-stage decomposition would compute slightly different F-values. Both analyses converge on the same qualitative conclusions.

\paragraph{Deviation from pre-registration}

\textbf{Deviations from the concept-note pre-registration disclosed here for transparency:}

\begin{itemize}
  \item \textbf{Closed-weight cells deferred.} M5, H2, H8 pre-registered as closed-weight; run as open-weight variants (M5 as Mistral, H2 as qwen3.6/phi-4/qwen-coder, H8 as gemma-4/qwen-coder/mistral). RQ4 correspondingly deferred to follow-up study.
  \item \textbf{Model family lineup.} The concept note listed qwen3.5:9B, llama3.3:70B; the swept lineup uses qwen3.6:27B, gemma4:26B, mistral-small3.2:24B. The change reflects the models available at sweep time (2026-06 to 2026-07).
  \item \textbf{Sample size asymmetric.} Concept note specified 150 per cell across all strata; swept as 150 mono, 75 hetero and null.
  \item \textbf{Held-out subsets partially swept.} The 25-sample HumanEval-Mutated subset was swept on H1 and M3 (Section~\ref{sec:results_contamination}). The complementary 25-problem LiveCodeBench post-cutoff subset was prepared per pre-registration but blocked at load time by a HuggingFace \texttt{datasets} library incompatibility; a full post-cutoff replication remains future work.
  \item \textbf{Human evaluation reported in addendum.} The 60-pair five-annotator study is reported in Section~\ref{sec:results_human_eval}, which post-dates the initial revision.
\end{itemize}

Every deviation is disclosed in this chapter and reflected in the source-controlled DOE table (\texttt{doe.py}), the pre-registered concept note (\texttt{chapter3\_concept\_note.md}), and the sweep TSV timestamps. No results were re-planned in light of the observed data.

\emph{End of Section~\ref{sec:discussion}.} Section~\ref{sec:conclusion} concludes.
