\section{Related Work}

This work sits at the intersection of five active research threads: (1) LLM-based unit-test generation, particularly under the mutation-testing evaluation paradigm; (2) LLM-based documentation generation from source code; (3) round-trip and consistency-based evaluation of language models; (4) LLM-as-judge protocols for semantic-equivalence assessment; and (5) multi-agent architectural specialisation in generative AI systems. We survey each thread in turn and position the present study.

\subsection{LLM-based unit test generation}

The dominant paradigm for automated unit-test generation prior to 2022 was \emph{search-based software testing} (SBST): tools such as EvoSuite for Java~(Fraser and Arcuri, 2011) and Pynguin for Python~(Lukasczyk and Fraser, 2022) treat test-suite synthesis as an optimisation problem, evolving candidate tests against a coverage- or mutation-based fitness function. These tools achieve high branch coverage on self-contained functions but suffer from the \emph{regression oracle problem} (assertions encode observed return values rather than specifications) and a \emph{per-function search-budget cost} that scales poorly to large codebases.

The advent of code-generation LLMs has shifted this landscape. \citet{Schafer2023LLMTest} demonstrated that GPT-4 can produce pytest-formatted test suites that read like hand-written tests and encode specifications drawn from docstrings or function signatures. \citet{Tufano2022LLMTest} benchmarked LLM-generated tests against SBST baselines under coverage metrics. \citet{Pan2023LLMSE} explored retrieval-augmented generation (RAG) variants for test generation on domain-specific codebases. These studies broadly converge on the finding that LLM-generated tests are competitive with SBST on branch coverage while requiring no per-function search budget.

The second companion paper of this series addresses three methodological gaps in this literature: (i) surface-level evaluation metrics (BLEU, ROUGE, syntactic validity) do not measure whether tests catch real defects; (ii) single-LLM studies do not surface interaction effects between LLM capability and RAG-method choice; and (iii) LLM-vs-SBST comparisons rarely appear on Python-native benchmarks with mutation-testing metrics. The second companion paper's cross-LLM $\times$ cross-method matrix (four methods $\times$ four LLMs on 300 HumanEval/MBPP functions) established the mutation kill rate as the SE-validated evaluation signal and produced the per-operator capability map (phi-4 on predicate reasoning; qwen-coder on comparison operators; the second companion paper's Table 13) that this paper's DOE takes as its per-stage strength oracle.

The present study extends the mutation-testing evaluation paradigm to the \emph{cross-stage} setting: mutation kill rate is used as the closure metric for Path~1 (Eq.~\ref{eq:killrate}), and the resulting per-cell kill rates are decomposed by mutation operator to test whether the second companion paper's per-operator capabilities transfer to the multi-stage setting (Section~4.4).

\subsection{LLM-based documentation generation}

Retrieval-augmented LLM docstring generation was surveyed by \citet{Liu2024RAGSurvey} and \citet{Zhang2023RAGCode}. The first companion paper of this series established the cross-LLM effect on docstring quality using a retrieval-augmented pipeline that grounds LLM outputs in framework documentation. That paper's finding --- that LLM choice matters for documentation generation under identical retrieval and prompting conditions --- provides the first plank of the thesis-level argument that method $\times$ model interactions in software-engineering pipelines cannot be recovered from single-LLM studies.

The present study uses docstring generation as the first stage of Path~1 ($L_{\textsf{spec}}(C) \to D'$) and as the recovered artefact in Path~3 ($L_{\textsf{spec}}(T') \to D'$). Docstring quality is evaluated via BERTScore F1 on Path~3 and, more importantly, via judge SLM equivalence rating (Section~3.5.4), because Section~4.6 and Section~5.2 show that BERTScore alone is inadequate as a semantic-equivalence signal on the docstring--docstring closure task.

\subsection{Round-trip and consistency-based evaluation of LLMs}

Three prior studies overlap with the round-trip framing adopted here. \citet{Allamanis2024RTC} introduced \emph{Round-Trip Correctness} (RTC), an unsupervised evaluation protocol for code LLMs that composes code $\to$ natural-language $\to$ code and measures whether the reconstruction is functionally equivalent to the original. RTC is a single-LLM protocol on closed-weight models (GPT-4, Codex) and uses functional test execution as the closure metric. \citet{Zhang2025CodingTriangle} introduced the \emph{Coding Triangle} framework, evaluating triangular consistency between code, specification, and tests on competitive-programming benchmarks. The Coding Triangle study is single-LLM self-consistency, uses accuracy on hidden test suites as the closure signal, and explicitly notes ``model mixtures can enhance'' as future work. \citet{Tian2024MBCT} proposed \emph{Mutation-Based Consistency Testing} (MBCT) --- but uses mutation as the \emph{attack} on the LLM's understanding (do mutations change the LLM's output?) rather than as the closure metric (do the LLM's tests detect defects that reference-suite mutations introduce?).

The present study advances this line of work along five orthogonal axes:

\begin{enumerate}
  \item \textbf{Heterogeneous multi-SLM closure} as a research variable. RTC, Coding Triangle, and MBCT are single-LLM protocols; we test whether cross-LLM stage assignments differ significantly from single-LLM baselines.
  \item \textbf{Mutation kill rate as the closure metric}, adopting the SE-validated defect-detection metric of the second companion paper rather than surface accuracy or BLEU/ROUGE.
  \item \textbf{Open-weight coverage} with pre-registered statistical rigour: six SLMs from five distinct families, all under 30~B parameters.
  \item \textbf{Per-stage bottleneck decomposition} via leave-one-stage-out ablations (N2, N3) and mixed-effects analysis (Eq.~\ref{eq:mixedlogit}) to identify which stage owns closure success.
  \item \textbf{Human validation of the closure metric} via the 60-pair stratified three-annotator study (Section~\ref{sec:human_eval_addendum}), directly quantifying the judge SLM's own agreement with human consensus ($\kappa_{\textsf{quad}} = 0.138$).
\end{enumerate}

\subsection{LLM-as-judge protocols}

\citet{Zheng2023LLMJudge} introduced the LLM-as-judge paradigm for evaluating LLM outputs, arguing that a sufficiently strong judge model can substitute for human evaluation on many tasks. Subsequent studies have documented judge-model biases including position bias, verbosity bias, and self-preference bias~(Panickssery et al., 2024; Wang et al., 2024). \citet{Fanous2025SycEval} introduced SycEval, a benchmark quantifying sycophancy in judge models under user pressure. The Multi-Agent Debate framework of \citet{Du2023MAD} extended judging to multi-agent ensembles, though \citet{Wynn2025DebateHarm} subsequently showed that multi-agent debate can be actively harmful when debaters share underlying RLHF biases.

The present study uses DeepSeek-R1:14B as an external single-model judge for semantic-equivalence rating on all three closure paths. DeepSeek-R1 is chosen deliberately because it belongs to a family (DeepSeek) disjoint from every pipeline SLM in the DOE, avoiding self-preference bias. Section~4.6 reports the Pearson correlation between judge ratings and automated metrics per path; Section~5.2 stratifies the disagreement by benchmark, revealing that judge--BERTScore disagreement on Path~3 flips direction across HumanEval and MBPP --- a structural signature that supports the more general claim that BERTScore is functioning as a surface-form similarity signal rather than a semantic-equivalence signal. The judge SLM's own validity is quantified in the human-evaluation addendum (Section~\ref{sec:human_eval_addendum}): against the majority-vote rating from three independent human annotators on the 60-pair pre-registered sample, DeepSeek-R1 attains $\kappa_{\textsf{quad}} = 0.138$ (linear-weighted $\kappa = 0.106$), placing it in Landis and Koch's ``slight'' band---a direct empirical justification for the strict-AND two-signal policy that pairs the judge with an SE-validated metric on every closure decision.

\subsection{Multi-agent architectural specialisation}

The design pattern of assigning distinct roles across multiple LLM instances --- rather than encoding all responsibilities in a single monolithic prompt --- has emerged as a productive research direction in adjacent problem domains. The Constitutional AI framework of \citet{Bai2022ConstitutionalAI} argues that principle-driven AI feedback can substitute for raw human preference labels in reinforcement learning. Concurrent work by \citet{Xiong2025DelphiAgent} on DelphiAgent introduces role-differentiated agents that reach consensus through iterative rounds of independent factuality judgements. \citet{Meier2025CausalDelphi} orchestrate multiple smaller specialised models for causal discovery via structured Delphi workflows.

The most directly parallel work in software engineering agents is the concurrent study of \citet{Anonymous2026EpistemicFortitude} on \emph{Epistemic Fortitude}: a two-agent (Primary + Arbiter) architecture that separates conversational fluency from truthfulness enforcement for sycophancy mitigation in software engineering agents. Their study demonstrates that architectural separation of Primary and Arbiter LLM roles produces statistically significant sycophancy reduction across four model families (Gemini~2.5, GPT-5.1, Claude Sonnet~4.5, Llama~3.3), with 49--77\% of the improvement attributable to \emph{architecture} rather than prompt content in low-baseline models.

The Epistemic Fortitude work is adjacent-but-distinct from the present study: they specialise LLMs by \emph{conversational role} (Primary vs.\ Arbiter within a single interaction turn) on sycophancy mitigation, while we specialise LLMs by \emph{artefact-generation stage} (spec vs.\ test vs.\ code across a multi-artefact pipeline) on round-trip closure of the docstring--test--code triangle. Both studies find that specialisation of role/stage across multiple LLMs beats monolithic single-LLM approaches on their respective task; both use LLM-as-judge protocols for scalable evaluation; and both find that per-model capability alone does not fully predict per-role/per-stage performance. We cite Epistemic Fortitude as an independent confirmation of the general architectural-specialisation-beats-monolith result on a distinct SE-adjacent task, and defer to that work for the sycophancy-mitigation problem it addresses.

\subsection{Positioning of this work}

The present study is, to our knowledge, the first cross-LLM $\times$ cross-stage empirical evaluation of round-trip closure in a software-engineering pipeline under a defect-detection-grounded validity metric. The five distinguishing contributions relative to the surveyed literature are:

\begin{itemize}
  \item A \textbf{pre-registered 20-cell fractional-factorial DOE} covering mono (6~cells), hetero (11~cells), and null (3~cells) strata, exhaustive at the mono stratum and hypothesis-driven at the hetero stratum. The DOE is committed to source before any sweep results are observed.
  \item \textbf{Mutation kill rate as the closure metric on Path~1}, with per-mutation-operator decomposition that both validates the pre-registered H1 hypothesis (phi-4 spec + qwen-coder test/code composition; Section~4.4) and exposes the phi-4-in-test-stage pathology at operator resolution (Section~5.3).
  \item An \textbf{external judge SLM protocol} paired with a strict two-signal validity policy (Algorithm~2), together with post-hoc categorical decomposition (Section~5.2) that turns Path~3's aggregate judge--metric non-correlation into a specific claim about BERTScore's surface-form character.
  \item \textbf{Per-stage bottleneck attribution} via the stage-contribution figure (Fig.~\ref{fig:stage_contribution}), the capability-preservation table (Table~\ref{tab:capability_preservation}), and the leave-one-stage-out null cells N2 and~N3.
  \item An \textbf{open, tested replication package} including 51-test unit coverage of the validity decision and test-filter algorithms.
\end{itemize}


\emph{End of Section~2 Related Work.} The remainder of the paper is organised as follows. Section~3 presents the methods, including the three closure paths, the validity decision, the DOE, and the statistical methodology. Section~4 reports the empirical results across the five research questions. Section~5 interprets these findings against RQ1--RQ5 and derives three design guidelines for practitioners. Section~6 details the threats to validity. Section~7 concludes with future research directions.
