\section{Conclusion}
\label{sec:conclusion}

We introduced the first cross-LLM, cross-stage empirical evaluation of round-trip closure in a software-engineering pipeline, evaluated under a pre-registered 20-cell design of experiments spanning six open-weight Small Language Models across the docstring--test--code triangle. The study answers the practitioner-facing question motivated in Section~1: \emph{when the pipeline is composed of multiple SLMs, does it matter which SLM sits at which stage?}

\subsection{Summary of contributions}

Five empirical contributions emerge from the 5{,}890-row sweep and the six post-hoc analyses (Section~5).

\begin{enumerate}
  \item \textbf{Pre-registered hypothesis validation at per-operator resolution.} The H1 cell (phi-4 as $L_{\textsf{spec}}$, qwen-coder as $L_{\textsf{test}}$ and $L_{\textsf{code}}$) was assigned in the pre-registered DOE based on the second companion paper's per-mutation-operator capability map. Section~\ref{sec:results_per_operator} shows that H1 dominates four of five mutation operators empirically, with the largest gain (arithmetic, $\Delta = +0.133$) matching the second companion paper's finding that phi-4 wins on predicate reasoning. Pre-registered stage compositions can be empirically validated by data at the operator level.

  \item \textbf{The phi-4-in-test-stage pathology.} Section~\ref{sec:discussion_rq3} documents a specific stage-model pairing --- phi-4 as $L_{\textsf{test}}$ --- that produces filter-failing test suites on approximately half of samples across three independent cells (M2 = $44\%$, H2 = $45\%$, H10 = $55\%$). The pathology localises at operator resolution to arithmetic and boundary mutations, the exact families phi-4 mono itself is competitive on. This is a cross-stage interaction effect invisible in single-stage evaluations.

  \item \textbf{BERTScore is a surface-form similarity signal on our data.} Section~\ref{sec:results_judge_metric_corr} reports that BERTScore F1 is statistically uncorrelated with the judge SLM's semantic-equivalence rating on Path~3 ($r_3 = 0.022$, $p = 0.359$, $n = 1{,}791$). Section~\ref{sec:discussion_rq2} stratifies this by benchmark and shows the disagreement flips direction: MBPP over-credits equivalence in $29\%$ of rows; HumanEval under-credits in $30\%$ of rows. This structural directional signature supports the specific claim that BERTScore behaves as a surface-form similarity signal rather than a semantic-equivalence signal on the docstring--docstring closure path.

  \item \textbf{The three-path framework carries non-redundant information.} Section~\ref{sec:results_contamination} reports the cell~$\times$~path interaction ANOVA (Eq.~\ref{eq:anova_interact}): $F(34, 4865) = 6.21$, $p < 0.001$, partial $\eta^2 = 0.042$. Cells respond \emph{differently} to different closure paths, not with a constant offset. The three-path framework of Section~3.2 is empirically justified rather than merely stipulated.

  \item \textbf{Capability preservation is not uniform.} Section~\ref{sec:results_stage_attribution} and Section~\ref{sec:discussion_capability} report that well-composed hetero cells (H11, H7, H9, H1) preserve $78$--$91\%$ of the strongest mono baseline's capability, while wrong-model-at-wrong-stage cells sacrifice materially (H10 preserves $34.9\%$ of Path~1 reference capability; H4 preserves $56.5\%$ of Path~2). Heterogeneity is not free; the stage assignment must match the demanded per-stage strength.
\end{enumerate}

\subsection{Design guidelines for practitioners}

Three guidelines follow from the empirical findings, targeted at practitioners assembling multi-SLM SE pipelines:

\begin{itemize}
  \item \textbf{G1: Place the strongest available model at the bottleneck stage.} The stage-attribution decomposition (Section~\ref{sec:results_stage_attribution}) and the per-operator table (Section~\ref{sec:results_per_operator}) both identify per-stage strengths that dominate closure success. Selecting stage models based on the pipeline's bottleneck --- rather than on aggregate model capability --- consistently produces higher closure rates.

  \item \textbf{G2: Match test-generation choice to the docstring style upstream, not to the model's general capability.} The phi-4-in-test-stage pathology (Section~\ref{sec:discussion_rq3}) shows that a model's test-generation capability \emph{de novo} does not transfer when the test model is downstream of a docstring produced by a different model. Stage models should be chosen jointly, not marginally.

  \item \textbf{G3: Use two-signal validity when reporting closure rates.} No single automated metric was found to correlate reliably with judge semantic equivalence on all three paths (Section~\ref{sec:results_judge_metric_corr}). The strict-AND validity policy of Algorithm~\ref{alg:validity} produces conservative closure counts with interpretable disagreement modes (Section~\ref{sec:discussion_rq2}). Single-signal validity risks systematic misestimation.
\end{itemize}

\subsection{Future work}

Four directions extend this study.

\begin{itemize}
  \item \textbf{Uniformly-drawn validation of the strict-AND policy.} The 60-pair human-evaluation sample of Section~\ref{sec:results_human_eval} is deliberately stratified toward metric--judge disagreement cases (20~frontier + 20~agreement + 20~disputed pairs). The FPR/FNR numbers therefore characterise the disagreement region, not a uniform draw from the full sweep. An immediate follow-up would draw a second 60-pair sample uniformly at random from the 5{,}890-row sweep and re-run the human annotation, allowing the strict-AND value proposition to be evaluated against a representative validity distribution. Estimated cost: five annotators $\times$ 60 pairs $\times$ $\sim$3 hours + API costs for the frontier-judge replay --- approximately one week of coordination.

  \item \textbf{RQ4 open-weight versus closed-weight ceiling comparison.} The pre-registered closed-weight cells (M5, H2, H8 in the original concept note) were deferred due to API-cost budget constraints. A follow-up study substituting Claude Sonnet~4.6 and GPT-4o-mini at the pre-registered slots would answer whether the closure-rate ceiling identified here for open-weight SLMs approaches, matches, or falls short of a closed-weight frontier ceiling. The pre-registered cells remain in \texttt{doe.py} with a \texttt{is\_deferred = True} marker; the DOE table is committed and unchanged.

  \item \textbf{Judge-model diversification.} Section~\ref{sec:results_human_eval} shows that four LLM judges (DeepSeek-R1:14B plus three frontier judges) cluster together in rating space at $\kappa_{\textsf{quad}} \approx 0.68$ but agree with human consensus only weakly ($\kappa_{\textsf{quad}} \leq 0.20$). Whether more architecturally diverse judges (a purpose-trained equivalence classifier, or a rule-based structural comparator) would break this cluster and improve human alignment remains open. Ensemble-of-judges designs~(Delphi-style consensus) are similarly unevaluated.

  \item \textbf{Full contamination replication.} The 25-sample HumanEval-Mutated held-out sweep on H1 and M3 (Section~\ref{sec:results_contamination}) rules out surface-form memorisation as a driver of the composition effects. A complementary post-training-cutoff replication (originally the LiveCodeBench-25 subset, blocked at load time by the removal of script-based dataset loaders from the current HuggingFace \texttt{datasets} library) would add date-based decontamination to the surface-form decontamination already reported. A recovered LiveCodeBench sweep across all six mono baselines and the four best hetero cells would extend the H1 dominance test to definitionally-uncontaminated problems.
\end{itemize}

\subsection{Position in the three-paper series}

This paper closes the empirical arc of the thesis on Small Language Models for software engineering. The first companion paper established that LLM choice matters materially for docstring generation. The second companion paper established that LLM choice matters materially for test generation, and that method $\times$ model interaction effects dominate the variance in test quality. The present study has now shown that the same interaction-effect phenomenon extends to the cross-stage setting: LLM choice interacts with \emph{stage assignment} across the docstring--test--code triangle. Specific stage-model pairings (phi-4 at $L_{\textsf{test}}$; llama-3.2 at $L_{\textsf{code}}$) exhibit pathologies invisible in single-stage evaluations; specific hetero cells (H1, H4, H11) achieve per-operator kill rates exceeding any single-model mono baseline; and the automated closure metric on the docstring--docstring path fails to carry the signal that a semantic-equivalence metric on that path should carry.

The single-sentence takeaway of the three-paper series: \emph{Small Language Models for software engineering are not interchangeable at any stage of a multi-stage pipeline, and the choice of which SLM sits at which stage is a first-class experimental variable whose effects are quantifiable, statistically significant, and reproducible from a pre-registered design of experiments.}
