%% ============================================================================
%% End matter — Elsevier mandatory closing sections.
%%
%% Order per Elsevier author guidelines:
%%   CRediT author statement
%%   Declaration of Competing Interest
%%   Acknowledgements
%%   Funding
%%   Ethics statement
%%   Data availability
%%   Reproducibility
%% ============================================================================

\section*{CRediT authorship contribution statement}

\textbf{Balaji Venktesh V.:} Conceptualization, Methodology, Software, Formal
analysis, Investigation, Data curation, Writing --- original draft, Writing ---
review \& editing, Visualization, Project administration.
\textbf{\PLACEHOLDER{Amsaprabhaa M.}:} Conceptualization, Methodology,
Supervision, Resources, Writing --- review \& editing, Project administration,
Funding acquisition.
\textbf{\PLACEHOLDER{Gireesh Sundaram}:} Methodology, Software, Validation,
Writing --- review \& editing.

\section*{Declaration of Competing Interest}

The authors declare that they have no known competing financial interests or
personal relationships that could have appeared to influence the work reported
in this paper.

\section*{Acknowledgements}

We thank \PLACEHOLDER{colleagues and reviewers} at Shiv Nadar University,
Chennai, for detailed feedback on the pre-registered design of experiments and
on the interim gate-check results. We are indebted to the anonymous reviewers
of the companion Chapters~1 and~2 submissions for feedback that shaped the
human-evaluation protocol and threats-to-validity structure adopted here. We
thank \PLACEHOLDER{annotator recruitment channel} for connecting us with the
three independent annotators who conducted the human-evaluation study, and we
thank the annotators themselves for the calibration and rating effort. This
work would not have been feasible without the open-weight releases from Meta,
Microsoft, Alibaba, Google, Mistral, and DeepSeek, and the maintainers of the
Ollama runtime, the LangChain / LangGraph ecosystem, and the BERTScore,
HumanEval, and MBPP benchmarks.

\section*{Funding}

This research received no specific grant from any funding agency in the public,
commercial, or not-for-profit sectors. The compute cost of the experimental
sweep was covered by the corresponding author's personal Colab Pro+
subscription. \PLACEHOLDER{Adjust to match actual funding once known.}

\section*{Ethics statement}

\paragraph{Data.} All datasets used in this study are publicly available prior
work: HumanEval~\citep{Chen2021HumanEval}, MBPP~\citep{Austin2021MBPP}, and
LiveCodeBench~\citep{Yang2025LiveCodeBench}. No new function-level data was
collected. The reference test suites, docstrings, and code accompanying these
benchmarks are used as-is.

\paragraph{Human participants.} The 60-pair human-evaluation study described in
Section~3 and reported in the pending addendum involves three independent human
annotators rating pairs of code artefacts (docstrings, code fragments, and test
suites) --- no personally-identifying data, biometric data, health data, or
demographic data is collected from annotators. Annotators are compensated at a
rate consistent with regional software-engineering hourly rates. The annotation
protocol was reviewed under the institutional research ethics framework at
\PLACEHOLDER{institution}; the review determined that the study was exempt from
full IRB review under the ``research involving benign educational tests or
observations of public behavior'' category. Annotator identities are not linked
to their ratings in any published artefact.

\paragraph{AI systems.} All experiments were conducted with open-weight
language models. No closed-weight models were used in the primary results. LLM
outputs were treated strictly as artefacts under evaluation --- the LLM systems
themselves are not the subject of ethical claims. The judge SLM (DeepSeek-R1:14B)
was used exclusively as an evaluation tool, not as a decision-maker for any
human-consequential decision.

\paragraph{Environmental impact.} The experimental sweep consumed approximately
22~GPU-days on Google Colab hardware, split between A100 and T4 instances.
Estimated CO$_2$-equivalent emissions:
\PLACEHOLDER{compute using MLCO2 estimator}. We disclose this as a matter of
methodological transparency and to enable readers to weight the environmental
cost against the study's scientific yield.

\section*{Data availability}

The complete replication package for this study is publicly available at
\url{https://github.com/balajivenky06/roundtrip-closure} under the MIT
licence. The repository contains: the full experimental sweep TSV
(\texttt{results/results\_roundtrip.tsv}, 5{,}890 measurement rows covering
all 20 cells of the pre-registered DOE); the 30-function pilot TSV; the
pre-registered DOE table (\texttt{doe.py}); the model lineup
(\texttt{config.py}); the three closure-path drivers (\texttt{closure\_paths.py});
the closure-validity decision (\texttt{closure\_decision.py}, Algorithm~2); the
test-filter validity gate (\texttt{closure\_metrics.filter\_tests\_with\_reason},
Algorithm~3); the mutation-testing engine inherited from Chapter~2
(\texttt{mutation\_testing.py}); the 51-test suite covering both algorithms
(\texttt{tests/test\_closure\_decision.py},
\texttt{tests/test\_filter\_tests\_reason.py}); the analysis pipeline
(\texttt{scripts/run\_analysis.py}); all six gap-closer scripts
(\texttt{scripts/build\_*.py}); the 11 LaTeX tables and 10 PNG figures used in
this manuscript, reproducible from the sweep TSV via a single command; and the
60-pair human-evaluation worksheet (blinded) and pair de-blinding index.

The pre-processed HumanEval, MBPP, LiveCodeBench, and HumanEval-Mutated datasets
are included in the \texttt{data/} directory as JSONL files. The five-layer
checkpointing infrastructure (LLM call cache, pytest subprocess cache,
BERTScore embedding cache, TSV row cache, JSONL decontamination cache)
preserves every intermediate artefact and enables exact re-execution of any
single path on any single sample without re-invoking the underlying SLMs.

\section*{Code and reproducibility}

The single command
\begin{verbatim}
python3 scripts/run_analysis.py --tsv results/results_roundtrip.tsv
\end{verbatim}
regenerates all 11 LaTeX tables and all 10 PNG figures from the sweep TSV. The
following commands regenerate the six gap-closer artefacts described in
Section~5:

\begin{itemize}
  \item G1 (threshold sensitivity):
        \texttt{python3 scripts/build\_threshold\_sensitivity.py}
  \item G2 (per-mutation-operator, requires LLM cache):
        \texttt{python3 scripts/regen\_per\_operator.py} followed by
        \texttt{python3 scripts/build\_per\_operator\_table.py}
  \item G3 (judge-metric disagreement):
        \texttt{python3 scripts/build\_disagreement\_decomposition.py}
  \item G4 (capability preservation):
        \texttt{python3 scripts/build\_capability\_preservation.py}
  \item G5 (path~$\times$~cell interaction ANOVA):
        \texttt{python3 scripts/build\_path\_cell\_anova.py}
  \item G6 (stage-contribution figure):
        \texttt{python3 scripts/build\_stage\_contribution\_figure.py}
\end{itemize}

Re-executing the full 20-cell experimental sweep from cold is possible via
\texttt{python3 scripts/run\_pilot.py} (30-function pilot) followed by
\texttt{python3 train\_roundtrip.py} per cell. Full-sweep wall-clock cost on
Colab Pro+ A100: approximately 15--22~hours; on T4: 60--90~hours. The
five-layer checkpointing means partial re-execution is inexpensive.

\section*{Preprint disclosure}

An earlier version of this manuscript was made available as a preprint at
\PLACEHOLDER{arXiv identifier or preprint URL} on \PLACEHOLDER{date}. This is
the peer-review version; substantive changes from the preprint are:
\PLACEHOLDER{list any changes prompted by feedback}.
