%% ============================================================================
%% Section 8 — Human-evaluation addendum
%% ============================================================================
\section{Human-evaluation addendum}
\label{sec:human_eval_addendum}

The 60-pair human evaluation described methodologically in
Section~\ref{sec:methods} has been completed with five annotators: three
from the author team (BV, GS, AM) and two recruited independently and
calibrated on the shared rubric (DR, RR). This section reports the
inter-rater and judge--human agreement statistics on the full five-way
sample, contrasts them with the three-way (author-team-only) numbers
that the initial revision reported, and interprets the implications for
the strict-AND two-signal validity policy that underpins this study's
closure decisions.

\subsection{Study conduct}
\label{sec:human_eval_conduct}

Five annotators rated the 60 pre-registered pairs on the 5-level rubric
described in Section~\ref{sec:methods}. Three (BV, GS, AM) are members
of the author team; two additional annotators (DR, RR) were recruited
independently of the pipeline runners and attended the shared calibration
session on the ten rubric anchor items. Sampling was as pre-registered:
20 pairs from the ``frontier'' bucket (judge and metric agree on
validity, drawn uniformly from cells with mid-range closure rates), 20
from the ``agreement'' bucket, and 20 from the ``disputed'' bucket.
Annotators saw pairs in a per-annotator deterministic order (SHA-256
seed on the annotator identifier) and were blinded to cell identity,
closure path, judge rating, and bucket assignment. All five annotators
completed the full 60 pairs and provided a written justification for
every rating (0 empty across all $60 \times 5 = 300$ ratings). Only
one revision was issued in total (by GS).

The rating distributions differ meaningfully between the author-team and
external annotators. All five annotators place the modal rating at~3
(``equivalent''), but external annotator DR uses the low end of the
scale more aggressively (six ratings at $0$ or $1$ versus $1$--$5$ for
the other four), and external annotator RR is more permissive at the
top of the scale (20~ratings at~$4$, similar to AM's~21 but higher than
BV/GS/DR at $11$--$16$). The external annotators are therefore not
outliers as much as they occupy a wider band on the ordinal scale than
the more clustered author-team ratings.

\subsection{Agreement results}
\label{sec:human_eval_agreement_results}

\input{../tables/tab_human_eval_agreement.tex}

Table~\ref{tab:human_eval_agreement} reports the primary agreement
statistics on the full five-way sample.

\textbf{Adding independent annotators materially lowers Krippendorff's
$\alpha$.} On the three author-team annotators alone (BV, GS, AM), the
ordinal $\alpha$ was $0.376$ --- itself below Krippendorff's
$\alpha \geq 0.667$ threshold for tentative inferential use but at least
weakly positive. Adding DR alone drops $\alpha$ to $0.244$; adding RR as
well drops it further to $\alpha = 0.193$. This monotonic decline is
consistent with the interpretation that the author-team-only $\alpha$
was inflated by shared calibration (all three share supervisory or
co-authorship relationships and had similar prior exposure to the
pipeline design); independent annotators, even after attending the same
calibration session, exercise the rubric differently. We adopt the
five-way $\alpha = 0.193$ as the more honest headline number and report
the three-way as a documented contrast.

\textbf{Pairwise Cohen's $\kappa$ shows the same pattern.} Pairwise
linear-weighted $\kappa$ within the author team ranges $0.170$--$0.338$;
between author-team and external annotators it drops to $-0.017$--$0.185$
(with BV-vs-RR marginally negative). Quadratic-weighted $\kappa$, which
penalises adjacent-category disagreements less harshly and is the more
common ordinal convention, follows the same ordering but at somewhat
higher values ($0.334$--$0.426$ within team; $0.009$--$0.263$
across team-vs-external). All within-team pairs are in the ``fair'' band
of the Landis and Koch conventions; team-vs-external pairs mostly fall
in ``slight'' to ``fair''.

\textbf{At coarser resolution the picture remains strong.} All ten
pairwise \emph{within-1} agreement rates (fraction of pairs where the
two ratings differ by at most one category on the 5-level scale) fall
in $[75.0\%, 91.7\%]$. Notably, team-vs-external pairs still achieve
$75\%$--$85\%$ within-1 agreement. Annotators converge on the rough
band of semantic equivalence while genuinely disagreeing at the fine
gradations of the top of the scale (whether a pair is ``identical'' or
merely ``equivalent'') and at the ``equivalent'' vs.\ ``approximately
equivalent'' boundary. This pattern is consistent with prior
human-evaluation studies of LLM outputs where 5-level ordinal scales
routinely produce sub-$0.5$ $\alpha$ values driven by
adjacent-category noise.

\textbf{Judge SLM tracks human consensus only weakly, even more so
against the five-way majority.} The DeepSeek-R1 judge SLM used
throughout this study attains
$\kappa_{\textsf{lin}} = 0.081$ and $\kappa_{\textsf{quad}} = 0.074$
against the five-way majority-vote human rating---both in Landis and
Koch's ``slight'' band, and materially weaker than the same judge
scored against the author-team-only majority
($\kappa_{\textsf{quad}} = 0.138$ at three annotators). At the coarser
within-1 resolution the judge does reach $91.7\%$ agreement with the
five-way majority, matching the pairwise inter-annotator within-1 rates,
but at the price of essentially no discriminative information at the
category-exact level.

\textbf{The ``disputed'' bucket has the highest five-way human
agreement, though at low absolute values.} Per-bucket exact-agreement
rates (all five annotators pick the same rating) are $5.0\%$
(``frontier''), $0.0\%$ (``agreement''), and $10.0\%$ (``disputed'').
While the absolute rates are low, the counter-intuitive ordering
observed in the three-way analysis persists at five: humans converge
most exactly on the pairs where the judge and metric could not agree.
The reading in which the automated-signal disagreement in the disputed
bucket arose from unambiguously wrong judge or metric calls that humans
identify jointly remains supported.

\subsection{Empirical value of the strict-AND policy}
\label{sec:human_eval_strict_and}

We recompute the false-closure ablation against the updated five-way
majority-vote human ground truth. Ground truth is defined as
majority-vote human rating $\geq 3$ (equivalent-or-better); positive-class
prevalence on this sample is $86.7\%$ (up from $78.3\%$ on the
three-way ground truth --- majority ties, broken toward the higher
rating following the pre-registered convention, more often land at
$\geq 3$ with five raters).
Table~\ref{tab:strict_and_ablation} reports the false-closure rate (FPR)
and false-reject rate (FNR) at the paper's default operating point
($\tau = 0$, $\rho = 3$) and at two more conservative metric thresholds.

\begin{table}[t]
  \centering
  \caption{False-closure and false-reject rates under three decision
    policies on the 60-pair five-annotator-majority sample. Ground truth:
    5-way majority human rating $\geq 3$ (positive-class prevalence
    $86.7\%$). Judge threshold $\rho = 3$ throughout. Columns report
    false-closure rate (FPR: fraction of human-negative pairs incorrectly
    labelled valid), false-reject rate (FNR: fraction of human-positive
    pairs incorrectly labelled invalid), and overall accuracy.}
  \label{tab:strict_and_ablation}
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    & \multicolumn{3}{c}{Metric-only}
    & \multicolumn{3}{c}{Judge-only}
    & \multicolumn{3}{c}{Strict-AND} \\
    \cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
    $\tau$
    & FPR & FNR & Acc & FPR & FNR & Acc & FPR & FNR & Acc \\
    \midrule
    $0.0$ & $87.5\%$ & $13.5\%$ & $76.7\%$ & $87.5\%$ & $5.8\%$ & $83.3\%$
          & $75.0\%$ & $19.2\%$ & $73.3\%$ \\
    $0.5$ & $50.0\%$ & $23.1\%$ & $73.3\%$ & $87.5\%$ & $5.8\%$ & $83.3\%$
          & $37.5\%$ & $28.8\%$ & $70.0\%$ \\
    $0.8$ & $37.5\%$ & $28.8\%$ & $70.0\%$ & $87.5\%$ & $5.8\%$ & $83.3\%$
          & $25.0\%$ & $32.7\%$ & $68.3\%$ \\
    \bottomrule
  \end{tabular}
\end{table}

Three empirical findings emerge from Table~\ref{tab:strict_and_ablation}.

\textbf{First: at every threshold combination, strict-AND achieves the
lowest false-closure rate of the three policies.} At the paper's
default $\tau = 0$, adding the judge to the metric reduces FPR from
$87.5\%$ to $75.0\%$---a $12.5$-percentage-point reduction. At
$\tau = 0.5$ the reduction is $12.5$~pp ($50.0\% \to 37.5\%$); at
$\tau = 0.8$ it is again $12.5$~pp ($37.5\% \to 25.0\%$). The exact
$12.5$~pp constancy across thresholds is a small-sample coincidence
(the reduction corresponds to a single-pair swap at $n = 8$ negatives)
but the direction is robust: the judge contributes usable signal in
the AND composition even at the low $\kappa_{\textsf{quad}} = 0.074$
against individual human ratings.

\textbf{Second: the reduction in false-closure is bought at the cost
of higher false-reject.} Adding the judge to metric-only raises FNR
by $5.8$--$5.7$--$3.9$~pp across $\tau \in \{0, 0.5, 0.8\}$. The
trade-off is favourable: strict-AND buys $2$--$3\times$ more FPR
reduction than the FNR rise it costs, so the policy is justified
under any loss function weighting false-closure at least as heavily
as false-reject.

\textbf{Third: judge-only at $\rho = 3$ has the best accuracy of any
single policy on this sample ($83.3\%$).} This is not evidence that the
judge alone is preferable in practice: at $\rho = 3$ the judge is
highly permissive on this sample, and its overall accuracy is dominated
by the elevated positive-class prevalence ($86.7\%$ under the 5-way
majority). Metric-only and strict-AND achieve their accuracy through
substantive discrimination; judge-only at $\rho = 3$ largely predicts
positive.

The strict-AND policy is therefore \emph{empirically supported}, not
merely \emph{motivated}, by the human-evaluation data. The point
estimate of a $12.5$~pp reduction in false-closure at the paper's
default operating point should nonetheless be read with a $n = 8$
negatives caveat: the $95\%$ Wilson confidence interval on the observed
FPR reduction spans roughly $\pm 25$~pp on this small negative-class
sample, so the effect is directionally established but the magnitude
is imprecise.

Two secondary implications:

\begin{enumerate}
  \item \textbf{The \texttt{false\_closure\_candidate} bucket in
        Section~\ref{sec:discussion} is likely upper-bounded rather
        than point-estimated.} A judge with $\kappa_{\textsf{quad}}
        = 0.074$ against human consensus will incorrectly flag some
        rows as ``judge disagrees with metric'' when a human would
        concur with the metric. The per-cell rates of
        \texttt{false\_closure\_candidate} reported in
        Section~\ref{sec:discussion} should therefore be treated as
        upper bounds on the true false-closure rate.

  \item \textbf{Metric-false-negative rows may partially reflect
        judge over-crediting rather than pure metric conservatism.}
        On the 60-pair sample the judge slightly over-rates relative
        to the 5-way majority human (mean judge~$= 3.28$ vs.\ mean
        human~$= 3.20$). Rows labelled \texttt{metric\_false\_negative}
        in Section~\ref{sec:discussion} (metric below $\tau$, judge
        above $\rho$) therefore mix two populations: cases where the
        metric is genuinely conservative on human-equivalent artefacts,
        and cases where the judge over-credits an artefact humans would
        rate lower. The BERTScore benchmark-flip finding of
        Section~\ref{sec:discussion} is unaffected in its comparative
        statement across benchmarks but should be read as a
        judge-relative---not a human-relative---signal.
\end{enumerate}

\subsection{Limitations of the addendum}
\label{sec:human_eval_limitations}

Three limitations bound the strength of the conclusions above.

First, $n = 60$ is the pre-registered sample size but is small for
inferring narrow confidence intervals on $\alpha$; the addendum reports
point estimates only. A larger sample would tighten the estimates but is
prohibitive at the human-evaluation rate.

Second, the annotator pool has \emph{partial} independence from the
author team: two of five annotators (DR, RR) are independent of the
pipeline runners, but three (BV, GS, AM) share supervisory or
co-authorship relationships. The direct effect of this partial
independence is quantified in Section~\ref{sec:human_eval_agreement_results}
(three-way $\alpha = 0.376$ collapses to five-way $\alpha = 0.193$ when
independent annotators are added). A fully independent replication with
three or more annotators recruited via a blinded third-party channel
remains future work; the direction of expected drift is now empirically
established.

Third, the 5-level ordinal rubric was chosen for compatibility with the
judge SLM's 0--4 output scale, but human annotators may find fewer
distinguishable levels (a 3-level ``different / approximate / equivalent''
rubric) easier to converge on. The high within-1 rates suggest that
collapsing the top two levels into a single ``equivalent'' category
would raise both $\alpha$ and pairwise $\kappa$ into the acceptable
band.
